Ideas cascading into keystrokes -- modelling writing hesitations as finite mixture process

### 2023-12-09 report/appendix_fit2data.tex
Reorganised manuscript

Model fit was assessed by testing to what extent simulations obtained
from the model fit the observed data. From each model we obtained 100
simulations (predictions). These can be found in Figure
\ref{fig:prediction}. The model predictions fitted the data
substantially better for the mixture models. The worse fit was observed
for the unimodal Gaussian model which predicted large amounts of
negative keystroke intervals which are implausible. Using a log-normal
distribution removed these negative keystroke intervals.

@@

\caption{Fit to data. Comparison of 100 simulated (predicted) sets of data to observed data summarised by model. For illustration the x-axis was truncated at 2 secs.}(\#fig:prediction)
\end{figure}
\end{appendix}

>>>>

Model fit was assessed by testing to what extent simulations obtained
from the model fit the observed data. From each model we obtained 100
simulations (predictions). These can be found in Figure
\ref{fig:prediction}. The model predictions fitted the data
substantially better for the mixture models. The worse fit was observed
for the single distribution Gaussian model which predicted large amounts
of negative keystroke intervals which are implausible. Using a
log-normal distribution removed these negative keystroke intervals.

@@

\caption{Fit to data with models (distribution type) above the plots. Comparison of 50 simulations to observed data summarised by model. For illustration the x-axis was truncated at 2 secs.}(\#fig:prediction)
\end{figure}
\end{appendix}

### 2023-12-09 report/appendix_genre_constr.Rmd
Reorganised manuscript

From the LIFT data, we compared the difference between genres, i.e. when writing an informative text as opposed to writing an argumentative text. The results are shown in Table \ref{tab:genreeffect}. Cellmeans and differences were average across writing topic. Evidence for a difference between genres was negligible.

@@

file <- list.files(path = "../stanout/lift", pattern = "mogbetaun.*.csv$",
                   full.names = T)

tmp <- read_csv(file) %>% 
  filter(param %in% c("beta", "delta", "theta")) %>% 
  mutate(across(value, ~ifelse(param == "theta", .*-1, value))) %>% 
  pivot_wider(names_from = genre, values_from = value) %>% 
  unnest(c(ARG, INF)) %>% 
  mutate(diff = ARG - INF) %>% 
  summarise(across(diff, list(mean = mean, 
                              lower = lower, 
                              upper = upper,
                              BF = BF), 
                   .names = "{.fn}"),
            .by = c(location, param)) %>% 
  mutate(across(mean, ~pmap_chr(list(., lower, upper, 2), PI)),
         across(BF, ~ifelse(.>100, "> 100", as.character(round(.,2)))),
         across(param, ~recode(., theta = "prob"))) %>% 
  left_join(cellmeans) %>% 
  arrange(param, location) %>% 
  select(param, 
         `Transition location` = location,
         `Argumentative` = ARG,
         `Informative` = INF, 
         "Difference" = mean, BF)

write_csv(tmp, "tables/genre_effect_cellmeans.csv")

apa_table(tmp[,-1],
          align = c("l", "r", "r", "r", "r"), 
          caption = "Mixture model estimates for genre effect. Cell means are shown for argumentative and informative texts in msecs for key-transitions, the slowdown for long transitions and the probability of hesitant transitions. The effect for genre is shown on log scale (for transition durations) and logit scale for probability of hesitant transitions. 95% PIs in brackets.",
          note = "PIs are probability intervals. BF is the evidence in favour of the alternative hypothesis over the null hypothesis.",
          longtable = T,
          font_size = "footnotesize",
          stub_indents = list(
            "Fluent transitions" = 1:3, 
            "Hesitation duration" = 4:6,
            "Hesitation probability" = 7:9))

>>>>

From the LIFT data, we compared the difference between genres, i.e. when writing an informative text as opposed to writing an argumentative text. The results are shown in Table \ref{tab:genreeffect}. Cellmeans and differences were average across writing topic. Evidence for a genre comparisons was negligible.

@@

file <- list.files(path = "../stanout/lift", pattern = "mogbetaco.*.csv$",
                   full.names = T)

tmp <- read_csv(file) %>% 
  filter(param %in% c("beta", "delta", "theta")) %>% 
  mutate(across(value, ~ifelse(param == "theta", .*-1, value)),
         across(genre, ~ifelse(is.na(genre) & param != "beta", "ARG", .)),
         across(genre, ~ifelse(. == "Inf", "INF", .))) %>% 
  pivot_wider(names_from = genre, values_from = value) %>% 
  unnest(c(ARG, INF)) %>% 
  mutate(diff = ARG - INF) %>% 
  summarise(across(diff, list(mean = mean, 
                              lower = lower, 
                              upper = upper,
                              BF = BF), 
                   .names = "{.fn}"),
            .by = c(location, param)) %>% 
  mutate(across(mean, ~pmap_chr(list(., lower, upper, 2), PI)),
         across(BF, ~ifelse(.>100, "> 100", as.character(round(.,2)))),
         across(param, ~recode(., theta = "prob"))) %>% 
  left_join(cellmeans) %>% 
  arrange(param, location) %>% 
  select(param, 
         `Transition location` = location,
         `Argumentative` = Arg,
         `Informative` = `Inf`, 
         "Difference" = mean, BF)

write_csv(tmp, "tables/genre_effect_cellmeans_constr.csv")

apa_table(tmp[,-1],
          align = c("l", "r", "r", "r", "r"), 
          caption = "Mixture model estimates for genre effect. Cell means are shown for argumentative and informative texts in msecs for the hesitation slowdown and the probability of hesitant transitions. The effect for genre is shown on log scale (for transition durations) and logit scale for probability of hesitant transitions. 95% PIs in brackets.",
          note = "PIs are probability intervals. BF is the evidence in favour of the alternative hypothesis over the null hypothesis.",
          longtable = T,
          font_size = "footnotesize",
          stub_indents = list(
            "Hesitation duration" = 1:3,
            "Hesitation probability" = 4:6))

### 2023-12-09 report/appendix_masking_constr.Rmd
Reorganised manuscript

Studies associated with two datasets (CATO, GUNNEXP2) investigated to what extent masking the previously written text affects keystroke behaviour. The mixture model results for the effect of masking is shown in Table \ref{tab:maskingeffect}. Key transition were longer for masked text than for unmasked text at before and within word transitions but not before sentences. Also dyslexics showed no masking difference at before word transitions. There was no indication of longer or more frequent pauses as a result of masked text.

@@

file <- list.files(path = "../stanout/gunnexp2", 
                   pattern = "mogbetaun.*.csv$",
                   full.names = T)

@@

# get posterior
file <- list.files(path = "../stanout/cato", 
                   pattern = "mogbetaun.+.csv$",
                   full.names = T)

@@

tmp <- bind_rows(d_gunn, d_cato) %>% 
  pivot_wider(names_from = mask, values_from = value) %>% 
  unnest(c(masked, unmasked)) %>% 
  mutate(diff = masked - unmasked) %>% 
  summarise(across(diff, list(mean = mean, 
                              lower = lower, 
                              upper = upper,
                              BF = BF), 
                   .names = "{.fn}"),
            .by = c(location, param, group, ds)) %>% 
  mutate(across(mean, ~pmap_chr(list(., lower, upper, 2), PI)),
         across(BF, ~ifelse(.>100, "> 100", as.character(round(.,2)))),
         across(param, ~recode(., theta = "prob")),
         across(ds, ~str_to_upper(.))) %>% 
  left_join(cellmeans_mask) %>% 
  arrange(param, location, ds, group) %>% 
  select(param, `Transition location` = location,
         `Dataset` = ds, Group = group, 
         Unmasked = unmasked,
         Masked = masked,
         "Difference" = mean, BF)

write_csv(tmp, "tables/masking_effect_cellmeans.csv")

apa_table(tmp[,-1],
          align = c("l", "l", "l", "r", "r", "r", "r"), 
          caption = "Mixture model estimates for masking effect. Cell means are shown for the masked and unmasked writing task in msecs for fluent key-transitions, the slowdown for long transitions and the probability of hesitant transitions. The effect for masking is shown on log scale (for transition durations) and logit scale for probability of hesitant transitions. 95% PIs in brackets.",
          note = "PIs are probability intervals. BF is the evidence in favour of the alternative hypothesis over the null hypothesis.",
          longtable = T,
          font_size = "footnotesize",
          stub_indents = list(
            "Fluent transitions" = 1:9, 
            "Hesitation duration" = 10:18,
            "Hesitattion probability" = 19:27) )

>>>>

Studies associated with two datasets (CATO, GUNNEXP2) investigated to what extent masking the previously written text affects keystroke behaviour. The mixture model results for the effect of masking is shown in Table \ref{tab:maskingeffect}. There was some evidence for longer before-sentence hesitation durations when the text was masked in non-dyslexic writers. Evidence for all other comparisons was negligible.

@@

file <- list.files(path = "../stanout/gunnexp2", 
                   pattern = "mogbetaco.*.csv$",
                   full.names = T)

@@

# get posterior
file <- list.files(path = "../stanout/cato", 
                   pattern = "mogbetaco.+.csv$",
                   full.names = T)

@@

tmp <- bind_rows(d_gunn, d_cato) %>% 
  pivot_wider(names_from = mask, values_from = value) %>% 
  unnest(c(masked, unmasked)) %>% 
  mutate(diff = masked - unmasked) %>% 
  summarise(across(diff, list(mean = mean, 
                              lower = lower, 
                              upper = upper,
                              BF = BF), 
                   .names = "{.fn}"),
            .by = c(location, param, group, ds)) %>% 
  mutate(across(mean, ~pmap_chr(list(., lower, upper, 2), PI)),
         across(BF, ~ifelse(.>100, "> 100", as.character(round(.,2)))),
         across(param, ~recode(., theta = "prob")),
         across(ds, ~str_to_upper(.))) %>% 
  left_join(cellmeans_mask, by = join_by(location, param, group, ds)) %>% 
  arrange(param, location, ds, group) %>% 
  select(param, `Transition location` = location,
         `Dataset` = ds, Group = group, 
         Unmasked = unmasked,
         Masked = masked,
         "Difference" = mean, BF)

write_csv(tmp, "tables/masking_effect_cellmeans_constr.csv")

apa_table(tmp[,-1],
          align = c("l", "l", "l", "r", "r", "r", "r"), 
          caption = "Mixture model estimates for masking effect. Cell means are shown for the masked and unmasked writing task in msecs for the slowdown for long transitions and the probability of hesitant transitions. The effect for masking is shown on log scale (for transition durations) and logit scale for probability of hesitant transitions. 95% PIs in brackets.",
          note = "PIs are probability intervals. BF is the evidence in favour of the alternative hypothesis over the null hypothesis.",
          longtable = T,
          font_size = "footnotesize",
          stub_indents = list(
            "Hesitation duration" = 1:9,
            "Hesitattion probability" = 10:18) )
